---
title: "wtr_analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r cars}
summary(cars)
```

## Load Libraries and Data

```{r pressure, echo=FALSE}
# Load Libraries
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lme4)
library(patchwork)
library(cowplot)


# Load Data 
df <- read.csv("wtr_data_clean.csv")

df_mod_count <- read_csv("wtr_model_pred_counts.csv")
df_mod_probs <- read_csv("wtr_model_pred_probs.csv")



```
# Demographics 
```{r}
df |> summary()

df %>%
  distinct(pid, gender) %>%
  count(gender)

```
# Data Visualizations


## Emprirical Data: Welfare Tradeoff Ratio by Donation Percentage and Knowledge Condition
```{r}
df %>%
  mutate(
    knows_endowment = as.logical(knows_endowment),
    ball_value = as.numeric(ball_value),
    condition = case_when(
      knows_endowment == FALSE ~ "ignores_endowment",
      knows_endowment == TRUE & ball_value == 20 ~ "knows_low",
      knows_endowment == TRUE & ball_value == 100 ~ "knows_high",
      TRUE ~ NA_character_
    )
  ) |> 
  ggplot(aes(x = donation_percentage, y = welfare_tradeoff_rating, color = condition)) +
  stat_summary(fun = mean, geom = "line", linewidth = 1) +
  stat_summary(fun = mean, geom = "point", linewidth = 3) +
  scale_color_manual(values = c(
    "ignores_endowment" = "#F8766D",
    "knows_high" = "#00BA38",
    "knows_low" = "#619CFF"
  )) +
  labs(
    x = "Donation Percentage",
    y = "WTR rating",
    color = "condition"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_line(color = "grey90"),
    legend.position = "bottom"
  )
```
## Model Predictions of WTR rating by Donation Amount and Knowledge Condition with model_pred_probs
```{r}
# Calculate expected value (weighted average)
ModelPredProbs_Amount_Knowledge <- df_mod_probs %>%
  mutate(
    condition = case_when(
      true_endowment == "unknown" ~ "ignores_endowment",
      true_endowment == "poor"  ~ "knows_low",
      true_endowment == "rich"  ~ "knows_high",
      TRUE ~ "other"  
    )
  ) %>%
  filter(condition != "other") %>%  
  
  # Calculate expected value: sum(rating * probability)
  rowwise() %>%
  mutate(
    expected_rating = (
      1 * `1` +
      2 * `2` +
      3 * `3` +
      4 * `4` +
      5 * `5` +
      6 * `6` +
      7 * `7`
    )
  ) %>%
  ungroup() %>%
  
  # Group and summarize
  group_by(donation_amount, condition) %>%
  summarise(
    mean_expected_rating = mean(expected_rating, na.rm = TRUE),
    se = sd(expected_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  
  # Create Plot
  ggplot(aes(x = donation_amount, y = mean_expected_rating, 
             color = condition, group = condition)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  # Add error bars (standard error)
  geom_errorbar(aes(ymin = mean_expected_rating - se, 
                    ymax = mean_expected_rating + se),
                width = 0.1, alpha = 0.5) +
  scale_color_manual(
    values = c(
      "ignores_endowment" = "#F8766D",
      "knows_high" = "#00BA38",
      "knows_low" = "#619CFF"
    ),
    labels = c(
      "ignores_endowment" = "Ignores Endowment",
      "knows_high" = "Knows High",
      "knows_low" = "Knows Low"
    )
  ) +
  labs(
    x = "Donation Amount",
    y = "Expected WTR Rating",
    color = "Condition"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_line(color = "grey90"),
    legend.position = "bottom"
  ) #+
 # scale_y_continuous(limits = c(1, 7), breaks = 1:7)

print(ModelPredProbs_Amount_Knowledge)
```
# Model Predictions by Donation Amount and WTR for Model Pred Counts 
```{r}
# Calculate expected value from counts
ModelPredCount_Amount_Knowledge <- df_mod_count %>%
  mutate(
    condition = case_when(
      true_endowment == "unknown" ~ "ignores_endowment",
      true_endowment == "poor"  ~ "knows_low",
      true_endowment == "rich"  ~ "knows_high",
      TRUE ~ "other"
    )
  ) %>%
  filter(condition != "other") %>%
  
  # Calculate expected rating from counts
  rowwise() %>%
  mutate(
    total_count = sum(c_across(`1`:`7`), na.rm = TRUE),
    expected_rating = ifelse(
      total_count > 0,
      (1 * `1` + 2 * `2` + 3 * `3` + 4 * `4` + 
       5 * `5` + 6 * `6` + 7 * `7`) / total_count,
      NA_real_
    )
  ) %>%
  ungroup() %>%
  group_by(donation_amount, condition) %>%
  summarise(
    mean_expected_rating = mean(expected_rating, na.rm = TRUE),
    se = sd(expected_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  
  # Create Plot
ggplot(aes(x = donation_amount, y = mean_expected_rating)) +
  geom_line(data = . %>% filter(condition == "ignores_endowment"),
            aes(group = condition), color = "#F8766D", size = 1.5, linetype = "solid") +
  geom_line(data = . %>% filter(condition == "knows_high"),
            aes(group = condition), color = "#00BA38", size = 1, linetype = "dashed") +
  geom_line(data = . %>% filter(condition == "knows_low"),
            aes(group = condition), color = "#619CFF", size = 1, linetype = "dashed") +
  geom_point(aes(color = condition), size = 3) +
  geom_errorbar(aes(ymin = mean_expected_rating - se, 
                    ymax = mean_expected_rating + se,
                    color = condition),
                width = 0.1, alpha = 0.5) +
  
  # Color scale for points and error bars
  scale_color_manual(
    values = c(
      "ignores_endowment" = "#F8766D",
      "knows_high" = "#00BA38",
      "knows_low" = "#619CFF"
    ),
    labels = c(
      "ignores_endowment" = "Ignores Endowment",
      "knows_high" = "Knows High",
      "knows_low" = "Knows Low"
    )
  ) +
  labs(
    x = "Donation Amount",
    y = "Expected WTR Rating",
    color = "Condition"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_line(color = "grey90"),
    legend.position = "bottom"
  ) +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7)

print(ModelPredCount_Amount_Knowledge)

```
#Model Predictions Plot by Donation Amount 
```{r}
model_newplot <-df_newpred %>% ggplot(aes(x=donation_amount, y=m.pred,
                   color=factor(condition,
                                levels=c('unknown', 'known_rich', 'known_poor')),
                   linetype=prior))+
  geom_point()+
  geom_line()+
  scale_color_discrete(name='condition')+
  scale_x_continuous(breaks=unique(df_newpred$donation_amount))+
  ylab('Mean Inferred WTR')+xlab('Donation Amount')+
  theme_minimal(base_size=14)
print(model_newplot)

```


# Final CogSci Model Plot (26.04.26)
```{r}
# New Model Prediction Plots (26.04.26)
# ---------- Define custom theme ----------
paper_theme <- theme(
  plot.title = element_text(size = 24, face = "bold", hjust = 0),
  plot.tag = element_text(size = 24, face = "bold"),

  axis.title = element_text(size = 24),
  axis.text.y  = element_text(size = 24),
  axis.text.x = element_text(size = 24),

  legend.title = element_text(size = 20, face = "bold"),
  legend.text  = element_text(size = 20),

  legend.position = "bottom",
  legend.box = "horizontal",
  legend.box.just = "center",

  panel.background = element_rect(fill = "white", color = NA),
  panel.grid.major = element_line(color = "grey80", size = 0.3),
  panel.grid.minor = element_line(color = "grey80", size = 0.1),

  plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
)

# ---------- Define color and linetype ----------
color_scale <- scale_color_manual(
  values = c(
    "ignores_endowment" = "#F8766D",
    "knows_high" = "#00BA38",
    "knows_low"  = "#619CFF"
  ),
  labels = c(
    "ignores_endowment" = "Unknown Endowment",
    "knows_high" = "Known High",
    "knows_low"  = "Known Low"
  )
)

linetype_scale <- scale_linetype_manual(
  values = c(
    "rich" = "solid",
    "poor" = "dotdash"
  ),
  labels = c(
    "rich" = "High Prior",
    "poor" = "Low Prior"
  )
)
#------------ Axis Compression Function --------------
compress_trans <- function() {
  scales::trans_new(
    name = "compress",
    transform = function(x) {
      ifelse(x <= 30, x, 30 + (x - 30)/3)  # Compress after 30 by factor of 3
    },
    inverse = function(x) {
      ifelse(x <= 30, x, 30 + (x - 30) * 3)
    }
  )
}
# --------------- Model Predictions Plot -------------
ModelPredCount_Amount_Knowledge_PriorNotCollapsed <- df_mod_count %>%
  mutate(
    condition = case_when(
      true_endowment == "unknown" ~ "ignores_endowment",
      true_endowment == "poor"    ~ "knows_low",
      true_endowment == "rich"    ~ "knows_high",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(condition)) %>%
  rowwise() %>%
  mutate(
    total_count = sum(c_across(`X1`:`X7`), na.rm = TRUE),
    expected_rating = ifelse(
      total_count > 0,
      (1*`X1` + 2*`X2` + 3*`X3` + 4*`X4` + 5*`X5` + 6*`X6` + 7*`X7`) / total_count,
      NA_real_
    )
  ) %>%
  ungroup() %>%
  group_by(donation_amount, condition, prior) %>%
  summarise(
    mean_expected_rating = mean(expected_rating, na.rm = TRUE),
    se = sd(expected_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = donation_amount,
    y = mean_expected_rating,
    color = condition,
    linetype = prior,
    group = interaction(condition, prior)
  )) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = mean_expected_rating - se, ymax = mean_expected_rating + se),
    width = 0.1, alpha = 0.5
  ) +
  color_scale +
  linetype_scale +
  labs(
    title = "Model Predictions",
    x = "Donation Amount",
    y = "Predicted WTR Rating",
    color = "Endowment",
    linetype = "Prior"
  ) +
  paper_theme +
    theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank()
  ) +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7)+
  scale_x_continuous(breaks = c(seq(0,30,5), 50, 75, 100),
                     trans = compress_trans())
    #scale_x_continuous(breaks = seq(0,100,5))

# --------------- Empirical Data Plot -----------------

#Summary Stats for Empirical Plot
Empirical_summary <- df %>%
  mutate(
    knows_endowment = as.logical(knows_endowment),
    ball_value = as.numeric(ball_value),
    condition = case_when(
      knows_endowment == FALSE ~ "ignores_endowment",
      knows_endowment == TRUE & ball_value == 20  ~ "knows_low",
      knows_endowment == TRUE & ball_value == 100 ~ "knows_high",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(condition), !is.na(prior)) %>%
  group_by(donation_amount, condition, prior) %>%
  summarise(
    mean_wtr = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se = sd(welfare_tradeoff_rating, na.rm = TRUE)/sqrt(n()),
    .groups = "drop"
  )

Empirical_Amount_Knowledge_PriorNotCollapsed <- Empirical_summary %>%
  ggplot(aes(
    x = donation_amount,
    y = mean_wtr,
    color = condition,
    linetype = prior,
    group = interaction(condition, prior)
  )) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_wtr - se, ymax = mean_wtr + se), width = 0.1, alpha = 0.5) +
  color_scale +
  linetype_scale +
  labs(
    title = "Empirical Data",
    x = "Donation Amount",
    y = "WTR Rating",
    color = "Endowment",
    linetype = "Prior"
  ) +
  paper_theme +
  theme(
    axis.title.y = element_blank()
  ) +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7) +
  scale_x_continuous(breaks = c(seq(0,30,5), 50, 75, 100),
                     trans = compress_trans())
  #scale_x_continuous(breaks = seq(0,100,5))

# --------- Remove legends from BOTH plots ----------
ModelPredCount_Amount_Knowledge_PriorNotCollapsed_no_leg <-
  ModelPredCount_Amount_Knowledge_PriorNotCollapsed + theme(legend.position = "none")

Empirical_Amount_Knowledge_PriorNotCollapsed_no_leg <-
  Empirical_Amount_Knowledge_PriorNotCollapsed + theme(legend.position = "none")

# -------- Extract shared legend from one plot -----------
shared_legend <- cowplot::get_legend(Empirical_Amount_Knowledge_PriorNotCollapsed)

# -------- Stack plots and add legend below -------------
combined_plot <- ModelPredCount_Amount_Knowledge_PriorNotCollapsed + 
                 Empirical_Amount_Knowledge_PriorNotCollapsed +
                 plot_layout(ncol = 1, heights = c(1,1), guides = "collect") & 
                 theme(
                   legend.position = "right",
                   legend.box = "vertical",
                   legend.box.just = "left",
                   legend.title = element_text(face = "bold", size = 20),
                   legend.text  = element_text(size = 20)
                 ) & 
  guides(
    linetype = guide_legend(order = 1,
                             override.aes = list(size = 2, shape = NA)),  
    color    = guide_legend(order = 2, override.aes = list(size = 3))
  )

# Add shared y-axis title 
final_plot <-
  cowplot::ggdraw(combined_plot) +
  cowplot::draw_label(
    "WTR Estimate",
    x = 0,
    y = 0.5,
    angle = 90,
    size = 24
  )
# Add margins to prevent cutoff
final_plot <- final_plot + 
  theme(plot.margin = margin(l = 30, r = 5, t = 0, b = 0))

print(final_plot)

```


## Correlation between model and empirical means for WTR 
```{r}
df_mod_count2 <- df_mod_count |>
  rowwise() |>
  mutate(
    mean_score =
      sum(c_across(`1`:`7`) * 1:7) /
      sum(c_across(`1`:`7`))
  ) |>
  ungroup() 

df_mod_count2 <- df_mod_count2 |>
  mutate(
    condition = true_endowment
  )
  

Empirical_means <- df %>%
  mutate(
    knows_endowment = as.logical(knows_endowment),
    ball_value = as.numeric(ball_value),
    condition = case_when(
      knows_endowment == FALSE ~ "unknown",
      knows_endowment == TRUE & ball_value == 20  ~ "poor",
      knows_endowment == TRUE & ball_value == 100 ~ "rich",
      TRUE ~ NA_character_
    )) 
Empirical_means <- Empirical_means |>
  select(
    prior,
    condition,
    donation_amount,
    welfare_tradeoff_rating
  )

Empirical_means2 <- Empirical_means |>
  group_by(prior,condition,donation_amount) |>
  summarise(
    mean_score = mean(welfare_tradeoff_rating, na.rm = T),
    .groups = "drop"
  )

#Combine data for Plot
combined_data <- bind_rows(
  Empirical_means2 %>% 
    select(prior, condition, donation_amount, mean_score) %>%  
    mutate(type = "participant"),
  
  df_mod_count2 %>% 
    mutate(type = "model")
)

#------------Correlation between Model and Participant Mean WTR ---------------
correlation_data <- combined_data %>%
  pivot_wider(
    id_cols = c(prior, condition, donation_amount),
    names_from = type,
    values_from = mean_score
  ) %>%
  # Remove rows where either is missing
  filter(!is.na(participant) & !is.na(model))

# Calculate correlation
correlation_result <- cor(
  correlation_data$participant, 
  correlation_data$model, 
  use = "complete.obs"
)
print(correlation_result)
```

## Mean WTR scatterplot for model and empirical data 
```{r}
scatterplot <- ggplot(correlation_data, aes(x = model, y = participant))+
  geom_abline(intercept = 0, slope = 1,
              linetype = "dashed", color = "gray60", linewidth = 1)+
  geom_point(aes(color = condition, shape = prior),
             size = 5, alpha = 0.8, stroke = 1.2)+
    labs(
    x = "Model Predictions (Mean WTR)",
    y = "Participant Data (Mean WTR)",
    title = "Correlation between Model Predictions\nand Participant Data",
    color = "Condition",
    shape = "Prior"
  ) +
    scale_color_manual(
    values = c(
      "unknown" = "#F8766D",
      "rich" = "#00BA38",
      "poor" = "#619CFF"
    ),
        labels = c(
      "unknown" = "Unknown Endowment",
      "rich" = "Known Rich",
      "poor" = "Known Poor"
    )
  ) +
    scale_shape_manual(
    values = c(
      "poor" = 4,      # X shape
      "rich" = 16      # Filled circle
    ),
        labels = c(
      "poor" = "Poor Prior",
      "rich" = "Rich Prior"
    )
  ) +
  # theme
  theme_minimal() +
  theme(
    # Title styling 
    plot.title = element_text(size = 32, face = "bold", hjust = 0.5, 
                              margin = margin(b = 15)),
    
    # Axis styling 
    axis.title = element_text(size = 24),
    axis.text = element_text(size = 20),
    
    # Legend styling 
    legend.title = element_text(size = 24, face = "bold"),
    legend.text = element_text(size = 20),
    
    # Legend position and arrangement
    legend.position = "bottom",
    legend.box = "vertical",       
    legend.box.just = "center",
    legend.spacing.y = unit(0.5, "cm"), 
    legend.box.spacing = unit(0.1, "cm"),
    
    # Panel styling 
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "grey80", size = 0.3),
    panel.grid.minor = element_line(color = "grey80", size = 0.1),
    
    # Plot margins
    plot.margin = margin(t = 20, r = 20, b = 15, l = 20)
  ) +
  
  # legend order 
  guides(
    shape = guide_legend(
      order = 1,
      title.position = "left",
      override.aes = list(size = 4, stroke = 1.2)
    ),
    color = guide_legend(
      order = 2,
      title.position = "left",
      override.aes = list(size = 4, shape = 16)  
    )
  ) 


print(scatterplot)



```
## Scatterplot of Empathy and WTR ratings 
```{r}
df_scatterplot_wtr_empathy <- df %>%
  mutate(
    knows_endowment = as.logical(knows_endowment),
    ball_value = as.numeric(ball_value),
    condition = case_when(
      knows_endowment == FALSE ~ "unknown",
      knows_endowment == TRUE & ball_value == 20  ~ "poor",
      knows_endowment == TRUE & ball_value == 100 ~ "rich",
      TRUE ~ NA_character_
    )
  )
empathy_wtr_scatterplot <- ggplot(df_scatterplot_wtr_empathy, aes(x = welfare_tradeoff_rating, y = empathy_rating)) +
geom_jitter(aes(color = condition),
            width = 0.15, height = 0.15,
            size = 3, alpha = 0.4, stroke = 1.2) +
  geom_smooth(method = "lm", se = TRUE, color = "gray40", linewidth = 1) +
    labs(
    x = "Welfare Tradeoff Rating",
    y = "Empathy Rating",
    title = "Correlation between Empathy Ratings\nand Welfare Tradeoff Ratings",
    color = "Condition"
  )+
  scale_color_manual(
    values = c(
      "unknown" = "#F8766D",
      "rich"    = "#00BA38",
      "poor"    = "#619CFF"
    ),
    labels = c(
      "unknown" = "Unknown Endowment",
      "rich"    = "Known Rich",
      "poor"    = "Known Poor"
    )
  ) +
  scale_shape_manual(
    values = c(
      "poor" = 4,   # X shape
      "rich" = 16   # Filled circle
    ),
    labels = c(
      "poor" = "Poor Prior",
      "rich" = "Rich Prior"
    )
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(size = 32, face = "bold", hjust = 0.5,
                                   margin = margin(b = 15)),
    axis.title      = element_text(size = 24),
    axis.text       = element_text(size = 20),
    legend.title    = element_text(size = 24, face = "bold"),
    legend.text     = element_text(size = 20),
    legend.position    = "bottom",
    legend.box         = "vertical",
    legend.box.just    = "center",
    legend.spacing.y   = unit(0.5, "cm"),
    legend.box.spacing = unit(0.1, "cm"),
    panel.background   = element_rect(fill = "white", color = NA),
    panel.grid.major   = element_line(color = "grey80", size = 0.3),
    panel.grid.minor   = element_line(color = "grey80", size = 0.1),
    plot.margin        = margin(t = 20, r = 20, b = 15, l = 20)
  ) +
  guides(
    shape = guide_legend(
      order = 1,
      title.position = "left",
      override.aes = list(size = 4, stroke = 1.2)
    ),
    color = guide_legend(
      order = 2,
      title.position = "left",
      override.aes = list(size = 4, shape = 16)
    )
  )

print(empathy_wtr_scatterplot)


```



# Statistical Models for Preregistered Hypotheses

#### Hypothesis 1: Participant judgments about the actor’s welfare trade-off ratio will be correlated with the model predictions about the actor’s welfare trade-off ratio

```{r}
## Add in correlation between model and participant judgments here
print(correlation_result)

updatedcor <- cor(df_newpred$m.wtr, df_newpred$m.pred)
print(updatedcor)
```

#### Hypothesis 2: Controlling for the amount of money shared, a prior belief about the giver having a high endowment will have a negative main effect on judgments about the welfare trade-off ratio

```{r}
mod2 <- lmerTest::lmer(welfare_tradeoff_rating ~ donation_amount + prior + (1 | pid), data = df  )
summary(mod2)
```

#### Hypothesis 3: Controlling for the amount of money shared, knowing the giver has a high endowment will have a negative main effect on judgments about the welfare trade-off ratio

```{r}
mod3 <- lmerTest::lmer(welfare_tradeoff_rating ~ donation_amount + knows_endowment + (1 | pid), data = df)
summary(mod3)

## How people marginalise between high and low endowment when they are in the unknown condition 
df_mod3 <- df %>%
  mutate(
  knows_endowment = as.logical(knows_endowment),
  ball_value = as.numeric(ball_value),
  condition = case_when(
    knows_endowment == FALSE ~ "ignores_endowment",
    knows_endowment == TRUE & ball_value == 20  ~ "knows_low",
    knows_endowment == TRUE & ball_value == 100 ~ "knows_high",
    TRUE ~ NA_character_
  )
)

## Is the unknown condition lower than the known_poor?
df_unknown_poor <- df_mod3 %>%
  filter(condition %in% c("ignores_endowment", "knows_low")) 

mod_unknown_poor <- lmerTest::lmer(welfare_tradeoff_rating ~ condition + (1 | pid), 
                         data = df_unknown_poor)

summary(mod_unknown_poor)

# Is the unknown condition higher than the known_rich?
df_unknown_rich <- df_mod3 %>%
  filter(condition %in% c("ignores_endowment", "knows_high") & 
         donation_amount < 25)

mod_unknown_rich <- lmerTest::lmer(welfare_tradeoff_rating ~ condition + (1 | pid), 
                         data = df_unknown_rich)

summary(mod_unknown_rich)

```


#### Hypothesis 4: Controlling for prior beliefs about endowment and the actual endowment, sharing money will have a positive main effect on judgments about the welfare trade-off ratio

```{r}
mod4 <- lmerTest::lmer(welfare_tradeoff_rating ~ prior + ball_value + donation_amount + (1 | pid), data = df)
summary(mod4)
```

#### Hypothesis 5: There will be an interaction between knowledge about the endowment and the amount of money the giver shares

```{r}
## Model with all donation amounts 
#df_mod5 <- df |> mutate(
 #knows_endowment = fct_relevel(factor(knows_endowment), "true" 
#))

mod5 <- lmerTest::lmer(welfare_tradeoff_rating ~ knows_endowment * donation_amount + (1 | pid), data = df)
summary(mod5)

## Model looking at the interaction specifically between $15 and $25 Pre-Registered 
df_subset <- df |>
  filter(
    donation_amount >= 15,
    donation_amount <= 25,
    prior == "poor",
    ball_value == "100"
  ) |>
  select(pid, knows_endowment, donation_amount, welfare_tradeoff_rating)

#df_subset <- df_subset |> mutate(
 # knows_endowment = fct_relevel(factor(knows_endowment), "true")
#)

mod5a <- lmerTest::lmer(welfare_tradeoff_rating ~ knows_endowment * donation_amount + (1 | pid), data = df_subset)
summary(mod5a)

## Exploratory model looking at the WTR estimates specifically between $20 and $25 in the unknown condition
df_subset2 <- df |>
  filter(
    donation_amount >= 20,
    donation_amount <= 25,
    knows_endowment == "false"
  ) |>
  select(pid, knows_endowment, donation_amount, welfare_tradeoff_rating)

df_subset2$donation_f <- factor(df_subset2$donation_amount)
#df_subset2$knows_endowment <- factor(df_subset2$knows_endowment)


mod5b <- lmerTest::lmer(welfare_tradeoff_rating ~ donation_amount + (1 | pid), data = df_subset2)
summary(mod5b)
```

#### Hypothesis 6: There will be a positive correlation between judgments about empathy and judgments about the welfare trade-off ratio

```{r}
cor.test(df$welfare_tradeoff_rating, df$empathy_rating)
```

# Exploratory Statistical Models

#### Exploratory 1: Controlling for donation amount (\$15) and knowledge about endowment (unknown), prior beliefs about endowment will have an effect on the welfare trade-off ratio

```{r}
df_exploratory_1 <- df %>%
  filter(
    donation_amount <= 20,
    knows_endowment == "false"
  )

mod_e1 <- lmerTest::lmer(welfare_tradeoff_rating ~ donation_amount + prior + (1 | pid), data = df_exploratory_1  )
summary(mod_e1)
```

#### Exploratory 2: Three-way interaction between Prior x Endowment Knowledge x Endowment

```{r}
df_exploratory_2 <- df %>%
  filter(
    donation_amount >= 15,
    donation_amount <= 25
  ) %>%
  select(pid, knows_endowment, donation_amount, welfare_tradeoff_rating, prior, ball_value)

# Ensure prior and knows_endowment are factors
df_exploratory_2$prior <- factor(df_exploratory_2$prior, levels = c("rich", "poor"))
df_exploratory_2$knows_endowment <- factor(df_exploratory_2$knows_endowment, levels = c("true", "false"))

mod_e2 <- lmerTest::lmer(welfare_tradeoff_rating ~ prior * knows_endowment * donation_amount + (1 | pid), data = df_exploratory_2)
summary(mod_e2)
```

### Exploratory 3: Holding donation percentage constant, in the knows endowment condition, do people make WTR inference driven by percentage
```{r}
df_exploratory_3 <-  df %>%
  filter(
    knows_endowment == "true"
  ) 

mod_e3 <- lmerTest::lmer(welfare_tradeoff_rating ~ donation_amount + donation_percentage + (1 | pid), df = df_exploratory_3)
```

### Exploratory 4: Calculating the Marginalisation between Known Conditions and Unknown Condition
```{r}
df_mod3 <- df %>%
  mutate(
  knows_endowment = as.logical(knows_endowment),
  ball_value = as.numeric(ball_value),
  condition = case_when(
    knows_endowment == FALSE ~ "ignores_endowment",
    knows_endowment == TRUE & ball_value == 20  ~ "knows_low",
    knows_endowment == TRUE & ball_value == 100 ~ "knows_high",
    TRUE ~ NA_character_
  ),
  condition = factor(condition, levels = c("knows_low", "ignores_endowment", "knows_high"))
)

mod3b <- lmer(welfare_tradeoff_rating ~ condition + (1 | pid), data = df_mod3)
summary(mod3b)

#Estimate the marginal means (group means)
library(emmeans)
emm <- emmeans(mod3b, ~ condition)

contrast(emm, method = list(
  marginalization = c(-0.5, 1, -0.5)
))
```


# Extra Data Visualisations 

## Welfare Trade-off Ratio by Donation Amount and Endowment Knowledge
```{r, fig.width=12, fig.height=7}
df %>%
  group_by(prior, donation_amount, knows_endowment) %>%
  summarise(
    mean_wtr = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se_wtr = sd(welfare_tradeoff_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
ggplot(
       aes(x = donation_amount,
           y = mean_wtr,
           color = knows_endowment,
           group = knows_endowment)) +
  geom_line(linewidth = 1) +         
  geom_point(size = 3) +             
  geom_errorbar(aes(ymin = mean_wtr - se_wtr,
                    ymax = mean_wtr + se_wtr),
                width = 2,
                linewidth = 0.8) +  
  facet_wrap(~ prior) +
  labs(
    x = "Donation Amount",
    y = "Mean Welfare Tradeoff Rating",
    color = "Knows Endowment",
    title = "Welfare Tradeoff Rating by Donation Amount\nLine Plot by Endowment Knowledge"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(size = 18),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Welfare Trade-off Ratio by Donation Percentage and Endowment Knowledge
```{r,fig.width=12, fig.height=7}
df %>%
  group_by(prior, donation_percentage, knows_endowment) %>%
  summarise(
    mean_wtr = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se_wtr = sd(welfare_tradeoff_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
ggplot(
       aes(x = donation_percentage,
           y = mean_wtr,
           color = knows_endowment,
           group = knows_endowment)) +
  geom_line(linewidth = 1) +        
  geom_point(size = 3) +             
  geom_errorbar(aes(ymin = mean_wtr - se_wtr,
                    ymax = mean_wtr + se_wtr),
                width = 2,
                linewidth = 0.8) +  
  facet_wrap(~ prior) +
  labs(
    x = "Donation Percent",
    y = "Mean Welfare Tradeoff Rating",
    color = "Knows Endowment",
    title = "Welfare Tradeoff Rating by Donation Percentage\nLine Plot by Endowment Knowledge"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(size = 18),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Welfare Trade-off Ratio by Donation Percentage and faceted by Prior and Endowment Knowledge

```{r}

# Convert ball value to factor 
df %>%
  mutate(ball_value = factor(ball_value)) |>
  group_by(prior, knows_endowment, donation_percentage, ball_value) %>%
  summarise(
    mean_wtr = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se_wtr = sd(welfare_tradeoff_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
# Plot
ggplot(
       aes(x = donation_percentage,
           y = mean_wtr,
           color = ball_value,
           group = ball_value)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_wtr - se_wtr,
                    ymax = mean_wtr + se_wtr),
                width = 2,
                linewidth = 0.8) +
  facet_grid(prior ~ knows_endowment) +  
  labs(
    x = "Donation Percentage",
    y = "Mean Welfare Tradeoff Rating",
    color = "Ball Drawn",
    title = "Welfare Tradeoff Rating by Donation Percentage\nFaceted by Prior and Endowment Knowledge"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

#### Same plot but with empathy instead of WTR ###

# Convert ball value to factor 
df %>%
  mutate(ball_value = factor(ball_value)) |>
# Summarize data with grouping including ball_drawn
  group_by(prior, knows_endowment, donation_percentage, ball_value) %>%
  summarise(
    mean_empathy = mean(empathy_rating, na.rm = TRUE),
    se_empathy = sd(empathy_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
#Plot
ggplot(
       aes(x = donation_percentage,
           y = mean_empathy,
           color = ball_value,
           group = ball_value)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_empathy - se_empathy,
                    ymax = mean_empathy + se_empathy),
                width = 2,
                linewidth = 0.8) +
  facet_grid(prior ~ knows_endowment) +   
  labs(
    x = "Donation Percentage",
    y = "Mean Empathy Rating",
    color = "Ball Drawn",
    title = " Empathy Rating by Donation Percentage\nFaceted by Prior and Endowment Knowledge"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```

## Welfare Trade-off Ratio by Donation Percentage and faceted by Prior and Endowment Knowledge

```{r}
# Convert ball value to factor 
df %>%
  mutate(ball_value = factor(ball_value)) |>
# Summarize data with grouping including ball_drawn
  group_by(prior, knows_endowment, donation_amount, ball_value) %>%
  summarise(
    mean_welfare_tradeoff = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se_welfare_tradeoff = sd(welfare_tradeoff_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
# Plot
ggplot(
       aes(x = donation_amount,
           y = mean_welfare_tradeoff,
           color = ball_value,
           group = ball_value)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_welfare_tradeoff - se_welfare_tradeoff,
                    ymax = mean_welfare_tradeoff + se_welfare_tradeoff),
                width = 2,
                linewidth = 0.8) +
  facet_grid(prior ~ knows_endowment) +   # facet rows = prior, cols = knows_endowment
  labs(
    x = "Donation Amount",
    y = "Mean Welfare Tradeoff Rating",
    color = "Ball Drawn",
    title = " Welfare Tradeoff Rating by Donation Amount\nFaceted by Prior and Endowment Knowledge"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

## Same Plot but with Empathy instead of WTR ## 

# Convert ball value to factor 
df %>%
  mutate(ball_value = factor(ball_value)) |>
# Summarize data with grouping including ball_drawn
  group_by(prior, knows_endowment, donation_amount, ball_value) %>%
  summarise(
    mean_empathy = mean(empathy_rating, na.rm = TRUE),
    se_empathy = sd(empathy_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
# Plot
ggplot(
       aes(x = donation_amount,
           y = mean_empathy,
           color = ball_value,
           group = ball_value)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_empathy - se_empathy,
                    ymax = mean_empathy + se_empathy),
                width = 2,
                linewidth = 0.8) +
  facet_grid(prior ~ knows_endowment) +   # facet rows = prior, cols = knows_endowment
  labs(
    x = "Donation Amount",
    y = "Mean Empathy Rating",
    color = "Ball Drawn",
    title = "Empathy Rating by Donation Amount\nFaceted by Prior and Endowment Knowledge"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Interaction between Endowment and Endowment Knowledge

```{r,fig.width=12, fig.height=7}
df %>%
  filter(
    donation_amount >= 15,
    donation_amount <= 25,
    prior == "poor",
    ball_value == "100"
  ) %>%
  select(pid, knows_endowment, donation_amount, welfare_tradeoff_rating) |>
  group_by(donation_amount, knows_endowment) %>%
  summarise(
    mean_wtr = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se_wtr = sd(welfare_tradeoff_rating, na.rm = TRUE)/sqrt(n()),
    .groups = "drop"
  ) |>
## Interaction Visualization
ggplot(aes(x = donation_amount, y = mean_wtr, color = knows_endowment)) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_wtr - se_wtr, ymax = mean_wtr + se_wtr),
                width = 0.5, linewidth = 0.8) +
  labs(
    x = "Donation Amount",
    y = "Mean Welfare Tradeoff Rating",
    color = "Knows Endowment",
    title = "Interaction of Endowment Knowledge and Donation Amount\n(Poor Prior, $100 Ball)"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5)
  )
```

## Welfare Tradeoff Ratio by Donation Amount and Prior faceted by Endowment
```{r}
#Convert Prior to Factor
df %>%
  mutate(prior = as.factor(prior)) |>
# Summarize data with grouping including ball_drawn
  group_by(prior, knows_endowment, donation_amount, ball_value) %>%
  summarise(
    mean_wtr = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se_wtr = sd(welfare_tradeoff_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) |>
#Plot
ggplot(
  aes(
    x = donation_amount,
    y = mean_wtr,
    color = knows_endowment,
    group = interaction(knows_endowment, prior),
    linetype = prior
  )
) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(
      ymin = mean_wtr - se_wtr,
      ymax = mean_wtr + se_wtr
    ),
    width = 2,
    linewidth = 0.8
  ) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  facet_wrap(~ ball_value) +
  labs(
    x = "Donation Amount",
    y = "Mean Welfare Tradeoff Rating",
    color = "Knowledge",
    linetype = "Prior",
    title = "Welfare Tradeoff Rating by Donation Amount\nFaceted by Prior and Endowment Knowledge"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    strip.text = element_text(size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

## Welfare Tradeoff Ratio by Donation Amount and Knowledge Condition
```{r}
Empirical_Amount_Knowledge <- df %>%
  mutate(
    knows_endowment = as.logical(knows_endowment),
    ball_value = as.numeric(ball_value),
    condition = case_when(
      knows_endowment == FALSE ~ "ignores_endowment",
      knows_endowment == TRUE & ball_value == 20 ~ "knows_low",
      knows_endowment == TRUE & ball_value == 100 ~ "knows_high",
      TRUE ~ NA_character_
    )
  ) |> 
  ggplot(aes(x = donation_amount, y = welfare_tradeoff_rating, color = condition)) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  scale_color_manual(values = c(
    "ignores_endowment" = "#F8766D",
    "knows_high" = "#00BA38",
    "knows_low" = "#619CFF"
  )) +
  labs(
    x = "Donation Amount",
    y = "WTR rating",
    color = "condition"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_line(color = "grey90"),
    legend.position = "bottom"
  )

print(Empirical_Amount_Knowledge)
```

## Welfare Tradeoff Ratio by Donation Percentage and Knowledge Condition
```{r}
df %>%
  mutate(
    knows_endowment = as.logical(knows_endowment),
    ball_value = as.numeric(ball_value),
    condition = case_when(
      knows_endowment == FALSE ~ "ignores_endowment",
      knows_endowment == TRUE & ball_value == 20 ~ "knows_low",
      knows_endowment == TRUE & ball_value == 100 ~ "knows_high",
      TRUE ~ NA_character_
    )
  ) |> 
  ggplot(aes(x = donation_percentage, y = welfare_tradeoff_rating, color = condition)) +
  stat_summary(fun = mean, geom = "line", size = 1) +
  stat_summary(fun = mean, geom = "point", size = 3) +
  scale_color_manual(values = c(
    "ignores_endowment" = "#F8766D",
    "knows_high" = "#00BA38",
    "knows_low" = "#619CFF"
  )) +
  labs(
    x = "Donation Percentage",
    y = "WTR rating",
    color = "condition"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_line(color = "grey90"),
    legend.position = "bottom"
  )
```

## Combined Plot with Model Predictions and Empriical Data for Expected WTR (Initial Plot - Not Paper Ready)
```{r}
# ---------- Define your custom theme ----------
paper_theme <- theme(
  plot.title = element_text(size = 32, face = "bold", hjust = 0),
  plot.tag = element_text(size = 28, face = "bold"),

  axis.title = element_text(size = 24),
  axis.text.y  = element_text(size = 20),
  axis.text.x = element_text(size = 20),

  legend.title = element_text(size = 24, face = "bold"),
  legend.text  = element_text(size = 20),

  legend.position = "bottom",
  legend.box = "horizontal",
  legend.box.just = "center",

  panel.background = element_rect(fill = "white", color = NA),
  panel.grid.major = element_line(color = "grey80", size = 0.3),
  panel.grid.minor = element_line(color = "grey80", size = 0.1),

  plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
)

# ---------- Define color and linetype scales once ----------
color_scale <- scale_color_manual(
  values = c(
    "ignores_endowment" = "#F8766D",
    "knows_high" = "#00BA38",
    "knows_low"  = "#619CFF"
  ),
  labels = c(
    "ignores_endowment" = "Unknown Endowment",
    "knows_high" = "Known Rich",
    "knows_low"  = "Known Poor"
  )
)

linetype_scale <- scale_linetype_manual(
  values = c(
    "rich" = "solid",
    "poor" = "dotdash"
  ),
  labels = c(
    "rich" = "High Prior",
    "poor" = "Low Prior"
  )
)
#------------ Axis Compression Function --------------
compress_trans <- function() {
  scales::trans_new(
    name = "compress",
    transform = function(x) {
      ifelse(x <= 30, x, 30 + (x - 30)/3)  # Compress after 30 by factor of 3
    },
    inverse = function(x) {
      ifelse(x <= 30, x, 30 + (x - 30) * 3)
    }
  )
}
# --------------- Model Predictions Plot -------------
ModelPredCount_Amount_Knowledge_PriorNotCollapsed <- df_mod_count %>%
  mutate(
    condition = case_when(
      true_endowment == "unknown" ~ "ignores_endowment",
      true_endowment == "poor"    ~ "knows_low",
      true_endowment == "rich"    ~ "knows_high",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(condition)) %>%
  rowwise() %>%
  mutate(
    total_count = sum(c_across(`1`:`7`), na.rm = TRUE),
    expected_rating = ifelse(
      total_count > 0,
      (1*`1` + 2*`2` + 3*`3` + 4*`4` + 5*`5` + 6*`6` + 7*`7`) / total_count,
      NA_real_
    )
  ) %>%
  ungroup() %>%
  group_by(donation_amount, condition, prior) %>%
  summarise(
    mean_expected_rating = mean(expected_rating, na.rm = TRUE),
    se = sd(expected_rating, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = donation_amount,
    y = mean_expected_rating,
    color = condition,
    linetype = prior,
    group = interaction(condition, prior)
  )) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(ymin = mean_expected_rating - se, ymax = mean_expected_rating + se),
    width = 0.1, alpha = 0.5
  ) +
  color_scale +
  linetype_scale +
  labs(
    title = "Model Predictions",
    x = "Donation Amount",
    y = "Predicted WTR Rating",
    color = "True Endowment",
    linetype = "Prior"
  ) +
  paper_theme +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7)+
  scale_x_continuous(breaks = c(seq(0,30,5), 50, 75, 100),
                     trans = compress_trans())
    #scale_x_continuous(breaks = seq(0,100,5))

# --------------- Empirical Data Plot -----------------

#Summary Stats for Empirical Plot
Empirical_summary <- df %>%
  mutate(
    knows_endowment = as.logical(knows_endowment),
    ball_value = as.numeric(ball_value),
    condition = case_when(
      knows_endowment == FALSE ~ "ignores_endowment",
      knows_endowment == TRUE & ball_value == 20  ~ "knows_low",
      knows_endowment == TRUE & ball_value == 100 ~ "knows_high",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(condition), !is.na(prior)) %>%
  group_by(donation_amount, condition, prior) %>%
  summarise(
    mean_wtr = mean(welfare_tradeoff_rating, na.rm = TRUE),
    se = sd(welfare_tradeoff_rating, na.rm = TRUE)/sqrt(n()),
    .groups = "drop"
  )

Empirical_Amount_Knowledge_PriorNotCollapsed <- Empirical_summary %>%
  ggplot(aes(
    x = donation_amount,
    y = mean_wtr,
    color = condition,
    linetype = prior,
    group = interaction(condition, prior)
  )) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_wtr - se, ymax = mean_wtr + se), width = 0.1, alpha = 0.5) +
  color_scale +
  linetype_scale +
  labs(
    title = "Empirical Data",
    x = "Donation Amount",
    y = "WTR Rating",
    color = "True Endowment",
    linetype = "Prior"
  ) +
  paper_theme +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7) +
  scale_x_continuous(breaks = c(seq(0,30,5), 50, 75, 100),
                     trans = compress_trans())
  #scale_x_continuous(breaks = seq(0,100,5))

# --------- Remove legends from BOTH plots ----------
ModelPredCount_Amount_Knowledge_PriorNotCollapsed_no_leg <-
  ModelPredCount_Amount_Knowledge_PriorNotCollapsed + theme(legend.position = "none")

Empirical_Amount_Knowledge_PriorNotCollapsed_no_leg <-
  Empirical_Amount_Knowledge_PriorNotCollapsed + theme(legend.position = "none")

# -------- Extract shared legend from one plot -----------
shared_legend <- cowplot::get_legend(Empirical_Amount_Knowledge_PriorNotCollapsed)

# -------- Stack plots and add legend below -------------
combined_plot <- ModelPredCount_Amount_Knowledge_PriorNotCollapsed +
                 Empirical_Amount_Knowledge_PriorNotCollapsed +
                 plot_layout(ncol = 1, heights = c(1,1), guides = "collect") &
                 theme(
                   legend.position = "bottom",
                   legend.box = "vertical",
                   legend.box.just = "center",
                   legend.title = element_text(face = "bold", size = 18),
                   legend.text  = element_text(size = 16)
                 ) &
  guides(
    linetype = guide_legend(order = 1,
                             override.aes = list(size = 2, shape = NA)),  # thicker lines in legend
    color    = guide_legend(order = 2, override.aes = list(size = 3))
  )

print(combined_plot)


```